---
title: "data_qc"
output: html_document
---
# Data quality control using R
<!-- source: https://obis.org/manual/processing/-->
```{r}
install.packages("devtools")

library(devtools)

install_github("iobis/robis")
install_github("iobis/obistools")
install.packages("ggplot2")
install_github("ropensci/finch")

library(robis)
library(obistools)
library(ggplot2)
library(finch)
library(obistools)
```

##### Checking points on land

The obistools package has a `check_onland()` function to check if coordinate pairs are located on land. By default this function uses a web service, but it can optionally work offline (although this is less accurate).

Fetch some Madrepora occurrences using robis:

```{r}
mad <- occurrence("Madrepora")
map_leaflet(mad)
```
Then run the `check_onland()` command. By default the function will return a data frame containing all records on land (another option is to return a data frame with errors):
```{r mad}
land <- check_onland(mad)

map_leaflet(land)
```
In some cases it makes sense to apply a buffer when checking for records on land. In this case we add a 1000 m buffer zone:

```{r mad}
land_buffer <- check_onland(mad, buffer = 1000)

map_leaflet(land_buffer)
```

As expected this returns less “wrong” records.

Now create a map showing all suspicious records, in orange by default but in red when they are suspicious even with the 1000 m buffer zone:
```{r land, land_buffer}
world <- map_data("world")

ggplot() +
 geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "#dddddd") +
 geom_point(data = land, aes(x = decimalLongitude, y = decimalLatitude), color = "#cc3300") +
 geom_point(data = land_buffer, aes(x = decimalLongitude, y = decimalLatitude), color = "#ff9900") + coord_fixed(1)
``` 

##### Taxon matching

The obistools package allows us to match taxa with the World Register of Marine Species directly from our R environment. To demonstrate this functionality, we are going to use the Reef Life Survey example dataset which is published on IPT here.

First, make sure the finch package is loaded, then read the Darwin Core Archive:

```{r}
archive <- dwca_read("http://ipt.iobis.org/obis-env/archive.do?r=rls-subset", read = TRUE) 
occurrence <- archive$data$occurrence.txt
```
Next, we can start the taxon matching procedure by passing our scientific names to the `match_taxa()` function:

```{r occurrence}
names <- match_taxa(occurrence$scientificName)
```

When the name matching has finished (this can take a while), a summary will be displayed indicating how many names were matches and how many need to be resolved manually:

Type info to see which names need manual action, y to start manual resolution, or n to skip manual resolution. After selecting y, several options will be presented for each name. Pick a number or press enter to skip the names:

After this procedure, you will end up with a data frame containing the matched name, the WoRMS LSID, and the type of match. Add the LSIDs to your source data as scientificNameID.

```{r occurrence, names}
occurrence$scientificNameID <- names$scientificNameID

View(names)

```

##### Check required fields

`check_fields()` will check if all OBIS required fields are present in an occurrence table and if any values are missing.This function returns a dataframe of errors (if any):

```{r}
data <- data.frame(
  occurrenceID = c("1", "2", "3"),
  scientificName = c("Abra alba", NA, ""),
  locality = c("North Sea", "English Channel", "Flemish Banks"),
  minimumDepthInMeters = c("10", "", "5")
)

check_fields(data)
```
## Plot points on a map

`plot_map()` will generate a `ggplot2` map of occurrence records, `plot_map_leaflet()` creates a Leaflet map.

```{r}
plot_map(abra, zoom = TRUE)

plot_map_leaflet(abra)
```

##### Checking depth values

The obistools package has a `check_depth()`, it uses the xylookup web service to identify which records have potentially invalid depths. Multiple checks are performed in this function:

    missing depth column (warning)
    empty depth column (warning)
    depth values that can't be converted to numbers (error)
    values that are larger than the depth value in the bathymetry layer, after applying the provided depthmargin (error)
    depth values that are negative for off shore points, after applying the provided shoremargin (error)
    minimum depth greater than maximum depth (error)

First download some occurrences from OBIS, then use `check_depth()` with a depthmargin of 10 meters, this will return all records where depth values are 10 meters or more below the bottom depth returned from the webservice:
```{r}
abrseg <- robis::occurrence("Abra segmentum")

problems <- check_depth(abrseg, depthmargin = 10)
```
To plot sample depth versus bottom depth, first use `lookup_xy()` to obtain bathymetry for our points:

```{r problems}
bathymetry <- lookup_xy(problems, shoredistance = FALSE, grids = TRUE, areas = FALSE)$bathymetry

plot(bathymetry, problems$maximumDepthInMeters) +
abline(0, 1, lty = 2)+
abline(10, 1, col = "red")

plot_map_leaflet(problems)

```
To print a report of the depth check:
```{r}
report <- check_depth(abra, report=T, depthmargin = 50)
head(report)
```
## Check outliers

`check_outliers_species` and `check_outliers_dataset` use the qc-service web service to identify which records are statistical outliers. For species outlier checks are performed for both environmental data (bathymetry, sea surface salinity and sea surface temperature) as well as spatially. Outliers are identified as all points that deviate more then six times the median absolute deviation (MAD) or three times the interquartile range (IQR) from the median. The `check_outliers_species` requires that the scientificNameID is filled in. When the species is already a known species in OBIS then the median, MAD and IQR from the known records are used for comparing the new species records. For datasets, only the spatial outliers are identified. Spatial outliers are detected based on the distance to the geographical centroid of all unique coordinates. The list in the extra field of the debug level output in the report provides all relevant statistics on which the outlier analysis is based. The `report` also gives an overview of these outliers. Outliers can be plotted with `plot_outliers(report)`

```{r abra}
report <- check_outliers_dataset(abra, report=T)
head(report)

plot_outliers(report)

?check_outliers_species
```
## Check eventID and parentEventID

`check_eventids()` checks if both `eventID()` and `parentEventID` fields are present in an event table, and if all `parentEventID`s have a corresponding `eventID`.

```{r}
data <- data.frame(
  eventID = c("a", "b", "c", "d", "e", "f"),
  parentEventID = c("", "", "a", "a", "bb", "b"),
  stringsAsFactors = FALSE
)
check_eventids(data)
```
## Check eventID in an extension

`check_extension_eventids()` checks if all `eventID`s in an extension have matching `eventID`s in the core table.

```{r}
event <- data.frame(
  eventID = c("cruise_1", "station_1", "station_2", "sample_1", "sample_2", "sample_3", "sample_4", "subsample_1", "subsample_2"),
  parentEventID = c(NA, "cruise_1", "cruise_1", "station_1", "station_1", "station_2", "station_2", "sample_3", "sample_3"),
  eventDate = c(NA, NA, NA, "2017-01-01", "2017-01-02", "2017-01-03", "2017-01-04", NA, NA),
  decimalLongitude = c(NA, 2.9, 4.7, NA, NA, NA, NA, NA, NA),
  decimalLatitude = c(NA, 54.1, 55.8, NA, NA, NA, NA, NA, NA),
  stringsAsFactors = FALSE
)

occurrence <- data.frame(
  eventID = c("sample_1", "sample_1", "sample_2", "sample_28", "sample_3", "sample_4", "subsample_1", "subsample_1"),
  scientificName = c("Abra alba", "Lanice conchilega", "Pectinaria koreni", "Nephtys hombergii", "Pectinaria koreni", "Amphiura filiformis", "Desmolaimus zeelandicus", "Aponema torosa"),
  stringsAsFactors = FALSE
)

check_extension_eventids(event, occurrence)
```

##### Working with OBIS-ENV-DATA

The obistools package provides some functionality to make it easier to work with OBIS-ENV-DATA (i.e. Event Core + Occurrence Extension + ExtendedMeasurementOrFact Extension).

First load a subset of the Reef Life Survey from IPT:

```{r}
archive <- dwca_read("http://ipt.iobis.org/obis-env/archive.do?r=rls-subset", read = TRUE)

event <- archive$data$event.txt
occurrence <- archive$data$occurrence.txt
mof <- archive$data$extendedmeasurementorfact.txt

```

Checking parentEventIDs in the Event Core

Use check_eventids() to check if all parentEventIDs in the Event Core have a matching eventID. In the example below, first the original data is checked, then an error is introduced, and then the data is checked again:

```{r event}
check_eventids(event)

event$parentEventID[event$eventID == "4000691_cryptic"] <- "dummy_1"

check_eventids(event)

```

Checking core record identifiers in the extension file

Use check_extension_eventids() to check if identifiers in an extension file have matching eventIDs in the Event Core file. Again, in the example first the correct data is checked and then an error is introduced. The function will return a table of errors (if there are any). The field parameter is the name of the identifier column in the extension file:

```{r event, occurrence}
check_extension_eventids(event, occurrence, field = "id")

occurrence$id[occurrence$occurrenceID == "48432"] <- "dummy_2"

check_extension_eventids(event, occurrence, field = "id")

```

##### Flattening occurrences

Sometimes it’s useful to have a flat occurrence table, i.e. a table of occurrences where all the information contained in the related events has been added. For example, all date and location information may be in the Event Core file and not in the Occurrence Extension, but for checking or analyzing your data you may want to have a table with both the date/location information and the scientific names. The field parameter is the column in the extension file which points to the core table:

 first go back to the clean version of the data, then look at the column names before and after flattening:


```{r archive}
event <- archive$data$event.txt

occurrence <- archive$data$occurrence.txt

mof <- archive$data$extendedmeasurementorfact.txt

flat <- flatten_occurrence(event, occurrence, field = "id")

names(occurrence)
names(flat)

```
## Calculate centroid and radius for WKT geometries

`calculate_centroid()` calculates a centroid and radius for WKT strings. This is useful for populating `decimalLongitude`, `decimalLatitude` and `coordinateUncertaintyInMeters`.

```{r}
wkt <- c(
  "POLYGON ((2.53784 51.12421, 2.99377 51.32031, 3.34534 51.39578, 2.82349 51.85614, 2.27417 51.69980, 2.53784 51.12421))",
  "POLYGON ((3.15582 42.23564, 3.13248 42.14202, 3.22037 42.11249, 3.26019 42.21530, 3.15582 42.23564))"
)

calculate_centroid(wkt)
```
## Map column names to Darwin Core terms

`map_fields()` changes column names using a terms map.

```{r}
data <- data.frame(
  id = c("cruise_1", "station_1", "station_2", "sample_1", "sample_2", "sample_3", "sample_4", "subsample_1", "subsample_2"),
  date = c(NA, NA, NA, "2017-01-01", "2017-01-02", "2017-01-03", "2017-01-04", NA, NA),
  locality = rep("North Sea", 9),
  lon = c(NA, 2.9, 4.7, NA, NA, NA, NA, NA, NA),
  lat = c(NA, 54.1, 55.8, NA, NA, NA, NA, NA, NA),
  stringsAsFactors = FALSE
)

mapping <- list(
  decimalLongitude = "lon",
  decimalLatitude = "lat",
  datasetName = "dataset",
  eventID = "id",
  eventDate = "date"
)

map_fields(data, mapping)
```
## Check eventDate

```{r}
data <- data.frame(eventDate = c(
  "2016",
  "2016-01",
  "2016-01-02",
  "2016-01-02 12",
  "2016-01-02 12:34",
  "2016-01-02 12:34:48",
  "2016-01-02T13:00:00+01:00",
  "2016-01-02T13:00:00+0100",
  "2016-01-02T13:00:00+01",
  "2016-01-02T13:00:00Z",
  "2016-01-02 13:00:00/2016-01-02 14:00:00",
  "2016-01-02 13:00:00/14:00:00",
  "2016-01-02/05",
  "2016-01-01 13u40",
  "2016/01/03",
  " ",
  NA
), stringsAsFactors = FALSE)

check_eventdate(data)
```

##### Visualizing dataset structure

The treeStucture() function in the obistools package generates a simplified event/occurrence tree showing the relationships between the different types (based on type and measurementType) of events and occurrences. Each node in the simplified tree is given a name based on the eventID or occurrenceID of one of the events of occurrences of that node type.

Note that an eventID column is required in the measurements table. In your dataset the extension column pointing to the event records may have another name, so make sure to add eventID.

```{r}
archive <- finch::dwca_read("http://ipt.iobis.org/obis-env/archive.do?r=nsbs&v=1.6", read = TRUE)
event <- archive$data$event.txt
occurrence <- archive$data$occurrence.txt
emof <- archive$data$extendedmeasurementorfact.txt
emof$eventID <- emof$id # adding eventID

tree <- treeStructure(event, occurrence, emof)
exportTree(tree, "tree.html")
```
## Data quality report 
<!--not working-->
`report()` generates a basic data quality report.
<!--
```{r}
report(abra)
?report
```
-->
## Lookup XY

`lookup_xy` returns basic spatial information for a given set of points. Currently three main things are returned: information from vector data (areas), information from rasters (grids) and the shoredistance.

```{r}
xydata <- lookup_xy(abra, shoredistance = TRUE, grids = TRUE, areas = TRUE)

head(xydata)

ggplot(xydata, aes(sssalinity, sstemperature, color=bathymetry)) +
    geom_point(alpha=0.5)
```
