---
title: "robis_visualizations"   
output: html_document
---
# Visualizing OBIS data
<!--source:https://obis.org/manual/visualization/-->

#### Introduction to the ggplot2 package

For making static maps in R, we are going to use the ggplot2 package. Getting some Southern sunfish occurrences from OBIS:

```{r}
install.packages("devtools")

library(devtools)

install_github("iobis/robis")
install.packages("ggplot2")
install.packages("dplyr")
install.packages("leaflet")
```

```{r}
library(robis)
molram <- occurrence("Mola ramsayi")

```

Now create a simple scatter plot using the occurrence coordinates. Use the ggplot() function to initialize a new plot, and geom_point() to create a scatter plot. The aes() function is used to create a mapping (aesthetic) between our data and the visual properties, which is then passed to the geom_point() function. Here decimalLongitude is used as the x coordinate and decimalLatitude as the y coordinate:

```{r molram}
library(ggplot2)
ggplot() + geom_point(data = molram, aes(x = decimalLongitude, y = decimalLatitude))

```

Let’s now add a world polygon to our map. ggplot2 provides a map_data() function to load maps of countries, continents or the entire world (this requires the maps package as well). Besides the world polygon, we also pass an aesthetic and a fill parameter to the geom_polygon() function. We will use coord_fixed() to make sure the axes have the same scale. You may want to pass a different value to this function if you are mapping areas close to the poles:

```{r molram}
library(maps)
world <- map_data("world")

ggplot() +
 geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "#dddddd") +
 geom_point(data = molram, aes(x = decimalLongitude, y = decimalLatitude))+
 coord_fixed(1)
 
```
 
 Let’s now zoom in a bit by passing xlim and ylim to coord_fixed:

```{r world molram}
ggplot() +
 geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "#dddddd") +
 geom_point(data = molram, aes(x = decimalLongitude, y = decimalLatitude)) +
 coord_fixed(1, xlim = c(0, 180), ylim = c(-60, 0))

```
 
 There are many ways to change color scales in ggplot, look into scale_color_brewer() for example:

```{r world molram}
ggplot() +
 geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "#dddddd") +
 geom_point(data = molram, aes(x = decimalLongitude, y = decimalLatitude, color = institutionCode)) +
 coord_fixed(1, xlim = c(0, 180), ylim = c(-60, 0)) +
 scale_color_brewer(palette = "Paired")

```
<!--this ones doesnt show anything much, and the aes is messy-->

The geom_bar() function can be used to create a graph. To try this, first fetch a bit more data from OBIS, and create a simple bar graph:

```{r}
dor <- occurrence("Doridoidea")

ggplot() +
 geom_bar(data = dor,stat='count', aes(x = year), binwidth = 2)
```

Binning the data by year doesn't always display nicely, so to give a better display, we will create a new column using dplyr mutate(), and do a bit of math to create 10 year bins for the data.

```{r dor}
library(dplyr)

binned_dor <- dor %>% mutate(yearBins=round(as.integer(year),digits=-1) )

                             
ggplot() +
 geom_bar(data = binned_dor,stat='count', aes(x = yearBins))

```

By adding fill to the aesthetic, we can color the bars based on the family, and using xlim() we can zoom in a bit:

```{r binned_dor}
ggplot() +
 geom_histogram(data = binned_dor, stat="count", aes(x = yearBins, fill = family)) +
 scale_fill_brewer(palette = "Spectral") +
 xlim(c(1950, 2017)) 
```
Moving to a new feature, in case you need to split up your graph based on one or more factors, you can use facet_grid(). For example:

```{r}
lag <- occurrence("Lagis")

lag_2 <- lag %>% filter(dataset_id %in% c("00032856-12cd-46ff-b9d5-ddeaecae3c95", "f0e5ebdf-0685-49b3-bca7-74e0835ec0dc"))

ggplot() +
 geom_histogram(data = lag_2,stat='count',  aes(x = year), binwidth = 2) +
 facet_grid(dataset_id~ species)

```
 
# Creating interactive maps using the leaflet package

The leaflet package is a wrapper around the popular Leaflet JavaScript library for interactive maps. Install the package as follows:

##### A simple map

Initialize a map with leaflet() and add the default OpenStreetMap basemap using addTiles(), to change the basemap, pick any of the tile providers here and pass the URL to addTiles():
:

```{r}
library(leaflet)

leaflet() %>% addTiles("https://server.arcgisonline.com/ArcGIS/rest/services/Ocean_Basemap/MapServer/tile/{z}/{y}/{x}")
```
Now fetch some data using the robis package, and add circle markers to the map using addCircleMarkers(). This function accepts lng, lat, as well as some styling patameters:


```{r}
abrseg <- occurrence("Abra segmentum")
popup <- paste0(abrseg$datasetName, "<br/>", abrseg$locality, "<br/>",abrseg$catalogNumber,"<br/><a href=\"http://www.iobis.org/explore/#/dataset/", abrseg$dataset_id, "\">OBIS dataset page</a>")


leaflet() %>%
  addTiles("https://cartodb-basemaps-{s}.global.ssl.fastly.net/light_all/{z}/{x}/{y}.png") %>%
  addCircleMarkers(popup = popup,lat = abrseg$decimalLatitude, lng = abrseg$decimalLongitude, radius = 3.5, weight = 0, fillOpacity = 1, fillColor = "#cc3300")

```
##### Multiple species

In the example below, data is retrieved and visualized for two cod species.
```{r}
pac <- occurrence("Gadus macrocephalus", startdate = '1993-01-01',enddate= '1995-01-01')

atl <- occurrence("Gadus morhua", startdate = '1993-01-01',enddate= '1995-01-01')

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(lat = pac$decimalLatitude, lng = pac$decimalLongitude, radius = 3.5, weight = 0, fillOpacity = 1, fillColor = "#ff0066") %>%
  addCircleMarkers(lat = atl$decimalLatitude, lng = atl$decimalLongitude, radius = 3.5, weight = 0, fillOpacity = 1, fillColor = "#0099cc")
```