---
title: "Mytilus edulis size data"
author: "Pieter Provoost"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---
<!--source: https://iobis.github.io/notebooks/size_data/-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

Run this cell if you are missing the required packages:

```{r message=FALSE, warning=FALSE, results='hide'}
install.packages("devtools")
library(devtools)
install_github("iobis/robis")
install.packages("dplyr")
install.packages("DT")
install.packages("stringr")
install.packages("ggplot2")
install.packages("sf")
install.packages("rnaturalearth")
```

First load the required packages:

```{r}
library(robis)
library(dplyr)
library(DT)
library(stringr)
library(ggplot2)
library(sf)
library(rnaturalearth)
```

Then use the `occurrence()` function with `mof = TRUE` to get all occurrences for Mytilus edulis, including the associated MeasurementOrFact data.

```{r message=FALSE, warning=FALSE, cache=TRUE}
occ <- occurrence("Mytilus edulis", mof = TRUE)
```

The `measurements()` function can be used to create a flat measurements table. Use the `fields` parameter to include occurrence fields in this table.

```{r message = FALSE, warning = FALSE}
mof <- measurements(occ, fields = c("decimalLongitude", "decimalLatitude", "collectionCode", "catalogNumber", "eventDate", "date_year", "dataset_id", "samplingProtocol", "samplingEffort"))

mof
```

This dataset includes a large variety of measurement types, let's create a list:

```{r message = FALSE, warning = FALSE}
types <- mof %>%
  group_by(measurementType, measurementTypeID, measurementUnit) %>%
  summarize(records = n(), datasets = length(unique(dataset_id))) %>%
  arrange(desc(records))

datatable(types)
```

There seem to be only two types of length measurements, both in mm. Let's select these measurements by their `measurementTypeID`, and convert the measurement values to numbers.

```{r message = FALSE, warning = FALSE}
size <- mof %>%
  filter(str_detect(measurementTypeID, "P01/current/OBSINDLX")) %>%
  mutate(measurementValue = as.numeric(measurementValue))
         
size
```

Now we can plot the size measurements by `collectionCode`:

```{r message = FALSE, warning = FALSE, fig.width=10, fig.height=7}
ggplot(size) +
  geom_jitter(aes(collectionCode, measurementValue, color = collectionCode), width = 0.1) +
  ylab("length (mm)")
```

And finally here are the measurements on a map:

```{r message = FALSE, warning = FALSE, fig.width=10, fig.height=7}
world <- ne_countries(scale = "large", returnclass = "sf")

ggplot() + 
  geom_sf(data = world) +
  geom_point(data = size, aes(decimalLongitude, decimalLatitude, size = measurementValue, color = collectionCode), pch = 21) +
  scale_size(range = c(0.1, 8)) +
  coord_sf(xlim = range(size$decimalLongitude), ylim = range(size$decimalLatitude)) +
  theme_void()
```