---
title: "Gobal record density map"
author: "Pieter Provoost"
output:
  html_document: default
editor_options:
  chunk_output_type: console
---
<!--source: https://iobis.github.io/notebooks/global_density_map/-->
<!-- data source: https://obis.org/manual/access/, Full OBIS export 2020-07-27-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

Install packages:

```{r}
install.packages("dplyr")
install.packages("stringr")
install.packages("sf")
install.packages("ggplot2")
install.packages("rnaturalearth")
install.packages("rnaturalearthdata")

```

Loading packages:

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(geosphere)
library(stringr)
library(readr)

world <- ne_countries(scale = "medium", returnclass = "sf")
filename <- "occurrence.csv"
```

## Data preparation

First create a global grid using `sf::st_make_grid`:

```{r, message = FALSE, warning = FALSE}
geom <- st_sfc(st_polygon(list(rbind(c(-180, -90), c(180, -90), c(180, 90), c(-180, 90), c(-180, -90)))), crs = 4326)
grid <- st_as_sf(st_make_grid(geom, cellsize = 1, square = TRUE))
```

This notebook uses a A CSV file with occurrence records exported from the database database. Download the latest export from [https://obis.org/manual/access](https://obis.org/manual/access). Only the coordinate columns are read using `readr::read_csv()`.

**Warning**: because intersection is very slow in `sf` I'm rounding coordinates here instead of doing an actual intersection with the grid. If you are not using a 1 degree rectangular grid, adjust accordingly.

```{r, message = FALSE, warning = FALSE}
df <- read_csv(filename, col_type = cols_only(decimallongitude = col_double(), decimallatitude = col_double()))

df_sf <- df %>%
  mutate(decimallongitude = floor(decimallongitude) + 0.5, decimallatitude = floor(decimallatitude) + 0.5) %>%
  group_by(decimallongitude, decimallatitude) %>%
  summarize(n = n()) %>%
  st_as_sf(coords = c("decimallongitude", "decimallatitude"), crs = 4326)
```

Now we can join the point layer with the grid.

```{r, message = FALSE, warning = FALSE}
grid_density <- grid %>%
  st_join(df_sf)
```

## Map

```{r, message = FALSE, warning = FALSE}
ggplot() + 
  geom_sf(data = grid_density, aes(fill = n), size = 0, color = NA) +
  geom_sf(data = world, fill = "#eeeeee", color = "#000000", size = 0.2) +
  scale_fill_viridis_c(
    option = "plasma",
    trans = "log10",
    na.value = "#eeeeee",
    name = "records"
  ) +
  coord_sf(crs = st_crs("ESRI:54030")) +
  theme_void()

ggsave(file = "records.png", width = 14, height = 10)
```
