---
title: "Assignment 5 - Identification is certain; confirmation is still necessary because of changed names"
output: html_document
author: Claude Nozères
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Identification is certain; confirmation is still necessary because of changed names

Finding new species in the Arctic is important, even if non-invasive, as indicators of environmental change.
Pacific sand lance was the name of species called Ammodytes hexapterus. Recently, it was genetically detected in the Arctic and thus an indicator of potential change (https://link.springer.com/article/10.1007/s00300-017-2141-0
). The identification is not in doubt. But does this need confirmation? Use select or filters to uncover the pattern of different Ammodytes records across the 3 oceans of Canada.

*Tip: review species names of Ammodytes hexapterus—common name is now Arctic sand lance. Pacific sand lance is now recognized as a new species, Ammodytes personatus. So old records of Pacific sand lanc as A. hexapterus are actually of A. personatus.*

*Tip: compare genetic results from BOLDsystems with occurrence records for Ammodytes spp.*

**Bonus: Ammodytes hexapterus is a historical name — at one time it was used for all, then was later split. Some records may have been updated using new information. Trace records (tip: compare from museum and fisheries surveys) for OriginalScientificName as may have been submitted to OBIS/GBIF.**

```{r }



```


