---
title: "Assignment 6 - Identification is uncertain, and names have also changed!"
output: html_document
author: Claude Nozères
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Identification is uncertain, and names have also changed!

A similar situation to the previous assignment may be explored with redfish (Sebastes spp.) in the Northwest Atlantic. Currently, 3 species are recognized—2 beaked species that are common (S. fasciatus, S. mentella) but very difficult to distinguish morphologically, and a third, unbeaked and less-common one (Sebastes norvegicus). A biologist may want to know if historical records of species of Sebastes vary in the distribution in Atlantic Canada.

What are some of the pitfalls of examining occurrence records over time and space for these species? Locate an example that illustrates the issue.

*Tip: all redfishes were once known as Sebastes marinus, for which the valid name is now Sebastes norvegicus. So past records show a common species was widespread, but then shrunk, when the 2 other species names came into use.*

```{r }



```


