---
title: "Assignment 8 - Detecting trends in vulnerable marine taxa - and uncovering data-deficient areas for analysis"
output: html_document
author: Claude Nozères
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Detecting trends in vulnerable marine taxa - and uncovering data-deficient areas for analysis

Much effort is underway to document and understand deepwater taxa such as corals and sponges. Sponges are difficult to identify, and were not always recorded in DFO groundfish trawl surveys. A number of fishing-excluded areas have been established. While research trawls continue in most if not all areas, a recent modelling exercise concluded these were unlikely to have significant impact (http://publications.gc.ca/collections/collection_2020/mpo-dfo/fs70-5/Fs70-5-2020-007-eng.pdf
). 

All DFO regions have published annual groundfish trawl survey data on OBIS (note: Quebec does not yet publish invertebrate data—currently being corrected for identifications). Are there areas, or taxa, of sponges that indicate trends in catch biomass for certain areas--or depths? Are trends evident in past five years, compared ot 10 years ago?

**Bonus: use species occurences cubes: https://www.biorxiv.org/content/10.1101/2020.03.23.983601v1**

```{r }



```


